Symbolic linearization of constrained multibody systems

### 2013-06-02 PetersonGedeHubbard2012.tex
Work on derivation of linearization section.
Linearization derivation section, results section, and parts of the discussion
section need to be merged. Examples still need to be added.

@@

@@

\begin{align}
\label{eq:quant_to_compute}
  \begin{array}{llcll}
    \tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 & \quad &
    \tilde{M}_{uqc} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_a \\
    \tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a & \quad &
    \tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_3 \\
    \tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 & \quad &
    \tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
    \tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 & \quad &
    \tilde{A}_{uqc} &\triangleq - \nabla_{\bm{q}} \bm{f}_a \\
    \tilde{A}_{uuc} &\triangleq - \nabla_{\bm{u}} \bm{f}_a & \quad &
    \tilde{A}_{uqd} &\triangleq - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
    \tilde{A}_{uud} &\triangleq - \nabla_{\bm{u}} \bm{f}_3 & \quad &
    \tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
  \end{array}
\end{align}

@@

%f_c:
%[r*cos(q1) + q5]
%f_v:
%[                r*u1*cos(q2) + u3]
%[-r*(u0*cos(q2) + u2*sin(q2)) + u4]
%[                r*u1*sin(q2) + u5]
%f_a:
%[                                -r*u1*sin(q2)*q2' + r*cos(q2)*u1' + u3']
%[-r*(-u0*sin(q2)*q2' + u2*cos(q2)*q2' + sin(q2)*u2' + cos(q2)*u0') + u4']
%[                                 r*u1*cos(q2)*q2' + r*sin(q2)*u1' + u5']
%f_0:
%[                                                                                      -sin(q2)*cos(q1)*q0' + cos(q2)*q1']
%[                                                                                                       sin(q1)*q0' + q2']
%[                                                                                       sin(q2)*q1' + cos(q1)*cos(q2)*q0']
%[(sin(q0)*cos(q2) + sin(q1)*sin(q2)*cos(q0))*q4' + (-sin(q0)*sin(q1)*sin(q2) + cos(q0)*cos(q2))*q3' - sin(q2)*cos(q1)*q5']
%[                                                                -sin(q0)*cos(q1)*q3' + sin(q1)*q5' + cos(q0)*cos(q1)*q4']
%[ (sin(q0)*sin(q2) - sin(q1)*cos(q0)*cos(q2))*q4' + (sin(q0)*sin(q1)*cos(q2) + sin(q2)*cos(q0))*q3' + cos(q1)*cos(q2)*q5']
%f_1:
%[-u0]
%[-u1]
%[-u2]
%[-u3]
%[-u4]
%[-u5]
%f_2:
%[                  -I*u0' - m*r*cos(q2)*u4']
%[-J*u1' + m*r*sin(q2)*u5' + m*r*cos(q2)*u3']
%[                  -I*u2' - m*r*sin(q2)*u4']
%f_3:
%[-I*u1*u2 + J*u1*u2 + g*m*r*sin(q1)*cos(q2) - m*r*(-u0*u5 + u2*u3)*cos(q2)]
%[                m*r*(u0*u4 - u1*u3)*sin(q2) + m*r*(u1*u5 - u2*u4)*cos(q2)]
%[ I*u0*u1 - J*u0*u1 + g*m*r*sin(q1)*sin(q2) - m*r*(-u0*u5 + u2*u3)*sin(q2)]

@@

\appendix
\section{Derivation of linearization procedure}
\label{sec:derivations}

This appendix outlines the derivation of Equation
(\ref{eq:state_space_constrained}. This equation can be used for simulation,
stability analysis, or it can be reformulated into the traditional state space
equations for a variety of tasks; these topics are discussed in Section
\ref{sec:discussion}.

Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
(\ref{eq:acceleration_expansion}),
and (\ref{eq:f2_plus_f3_expansion}) to zero (as in Table
\ref{table:assumptions}), and introducing the following definitions
enables the unconstrained linear state space equations to be written as
\begin{align}
  \label{eq:state_space_unconstrained}
  \left[
    \begin{array}{cc}
      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
      \tilde{M}_{uqd} & \tilde{M}_{uud}
    \end{array}
    \right]
    \left[
      \begin{array}{c}
        \delta \bm{\dot{q}} \\
        \delta \bm{\dot{u}}
      \end{array}
    \right]
   &=
   \left[
     \begin{array}{cc}
       \tilde{A}_{qq} & \tilde{A}_{qu} \\
       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
       \tilde{A}_{uqd} & \tilde{A}_{uud}
     \end{array}
   \right]
    \left[
      \begin{array}{c}
        \delta \bm{q} \\
        \delta \bm{u}
      \end{array}
    \right]
    +
    \left[
      \begin{array}{c}
        \bm{0}_{(n + m) \times s} \\
        \tilde{B}_{u}
      \end{array}
    \right]
    \delta \bm{r}
\end{align}
Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
+ o$, yet only $p = n - l + o - m$ of these quantities are independent.  To
address this issue, a smaller set of independent coordinates and speeds must be
selected. To this end, consider the following partitioning of the generalized
coordinates and generalized speeds:
\begin{equation*}
  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
      \qquad\qquad
  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
\end{equation*}
where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
are invertible permutation matrices which map an ordering which has the
independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities,
($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the original
ordering of the coordinates and speeds.  We use the notation $P_{qi}$ and
$P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$, respectively;
similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while $P_{ud}$ is the
last $m$ columns of $P_u$.

% Not sure which bibliography style we are supposed to use
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliographystyle{spbasic}      % basic style, author-year citations
\bibliography{references}   % name your BibTeX data base
\bibliographystyle{spmpsci}      % mathematics and physical sciences
\end{document}

>>>>

It is critical to note that just taking the gradient of (\ref{eq:F_Fstar_f_a})
with respect to $\mathbf{u}_i$ does not give the correct linear equations. The
reasons for this will be covered later.

@@

\section{Derivation of linearization procedure}
\label{sec:derivations}

In response to the need we have demonstrated, we present a solution to this
constrained, linearization problem.
There is a reason that the equations generated by taking the gradient with
respect to the generalized speeds are incorrect.
When finding the equations of motion for a constrained system, some generalized
coordinates and speeds are classified as dependent speeds, meaning their
behavior is determined by the other coordinates and speeds.
These terms should in fact be written as:
\begin{align}
\label{eq:q_d_redefined}
\mathbf{q}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t) \\
\label{eq:u_d_redefined}
\mathbf{u}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t)
\end{align}
Then, upon taking the derivative of each of the terms with respect to an
independent speed, the dependent speed would need to be solved for and its
derivative found.
No previous authors appear to have made clear that this distinction should be
made when using Kane's method.

What we present addresses this issue.
Rather then substituting in the solutions to the dependent speeds though, we
use knowledge of the structure of the equations and constraints (as defined in
Table \ref{table:assumptions}) to generate linear equations.

Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
(\ref{eq:acceleration_expansion}),
and (\ref{eq:f2_plus_f3_expansion}) to zero (as in Table
\ref{table:assumptions}), and introducing the following definitions
\begin{align}
\label{eq:quant_to_compute}
  \begin{array}{llcll}
    \tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 & \quad &
    \tilde{M}_{uqc} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_a \\
    \tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a & \quad &
    \tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_3 \\
    \tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 & \quad &
    \tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
    \tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 & \quad &
    \tilde{A}_{uqc} &\triangleq - \nabla_{\bm{q}} \bm{f}_a \\
    \tilde{A}_{uuc} &\triangleq - \nabla_{\bm{u}} \bm{f}_a & \quad &
    \tilde{A}_{uqd} &\triangleq - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
    \tilde{A}_{uud} &\triangleq - \nabla_{\bm{u}} \bm{f}_3 & \quad &
    \tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
  \end{array}
\end{align}
enables the unconstrained linear state space equations to be written as
\begin{align}
  \label{eq:state_space_unconstrained}
  \left[
    \begin{array}{cc}
      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
      \tilde{M}_{uqd} & \tilde{M}_{uud}
    \end{array}
    \right]
    \left[
      \begin{array}{c}
        \delta \bm{\dot{q}} \\
        \delta \bm{\dot{u}}
      \end{array}
    \right]
   &=
   \left[
     \begin{array}{cc}
       \tilde{A}_{qq} & \tilde{A}_{qu} \\
       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
       \tilde{A}_{uqd} & \tilde{A}_{uud}
     \end{array}
   \right]
    \left[
      \begin{array}{c}
        \delta \bm{q} \\
        \delta \bm{u}
      \end{array}
    \right]
    +
    \left[
      \begin{array}{c}
        \bm{0}_{(n + m) \times s} \\
        \tilde{B}_{u}
      \end{array}
    \right]
    \delta \bm{r}
\end{align}
Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
+ o$, yet only $p = n - l + o - m$ of these quantities are independent.  To
address this issue, a smaller set of independent coordinates and speeds must be
selected. To this end, consider the following partitioning of the generalized
coordinates and generalized speeds:
\begin{equation*}
  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
      \qquad\qquad
  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
\end{equation*}
where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
are invertible permutation matrices which map an ordering which has the
independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities,
($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the original
ordering of the coordinates and speeds.  We use the notation $P_{qi}$ and
$P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$, respectively;
similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while $P_{ud}$ is the
last $m$ columns of $P_u$.

@@

@@

@@

\appendix
% Not sure which bibliography style we are supposed to use
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliographystyle{spbasic}      % basic style, author-year citations
\bibliography{references}   % name your BibTeX data base
\bibliographystyle{spmpsci}      % mathematics and physical sciences
\end{document}
